Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-31 article/02_Materials_and_Methods.Rmd
more phrasing and minor improvements

```{r dacta, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_A.RData")
distance_and_correlation_table_A %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Mean|Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
  ) %>%
knitr::kable(
  "latex", 
  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance. Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B). $p$-values below $0.1$ are printed bold.", 
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "A" = 3, "B" = 3),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

```{r dactb, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_B.RData")
distance_and_correlation_table_B %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black")))
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1))
  ) %>%
knitr::kable(
  "latex",
  caption = "Like Table \\ref{tab:dacta}. \\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering and correcting for spatial distance as a second independent variable.",
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "variable correlation" = 2, "variable correlation with spatial correction" = 2)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "C" = 2, "D" = 2),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

>>>>

```{r dacta, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_A.RData")
distance_and_correlation_table_A %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Mean|Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
  ) %>%
knitr::kable(
  "latex", 
  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance. Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B). $p$-values below $0.1$ are printed bold.", 
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "type distance & spatial distance" = 3, "construction distance & spatial distance" = 3)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "A" = 3, "B" = 3),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

```{r dactb, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_B.RData")
distance_and_correlation_table_B %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black")))
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1))
  ) %>%
knitr::kable(
  "latex",
  caption = "Like Table \\ref{tab:dacta}. \\textbf{C} shows the Mantel test results for the distance networks based on the two variables if compared directly, \\textbf{D} the results of the respective Partial Mantel test considering and correcting for spatial distance as a second independent variable.",
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "type & construction distance" = 2, "type & construction & spatial distance" = 2)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "C" = 2, "D" = 2),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

### 2019-03-31 article/03_Results.Rmd
more phrasing and minor improvements

The overall picture is that during the Bronze Age in Europe there was a trend away from body burial towards cremation. Around 2200 calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany. In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions. These contexts could be discussed as areas of origin of the later omnipresent phenomenon. While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained relevance there in the Middle Bronze Age, their share increased steadily in England. Inhumation and cremation were balanced here for a long time. The situation was similar in Southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age. In Poland and northern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age. In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age. In both contexts inhumation continued to play an important role.

@@

Table \ref{tab:dactb} shows that the distance networks of *burial type* and *burial construction* are not equivalent. There is also no significant negative correlation. The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity of developments in these regions concerning the dominance of flat or mound graves. Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantle test is almost undercut, thus suggesting a statistically significant relationship. This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced.

@@

For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC. In the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to see the same developments. In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed. The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- although not significantly. The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.

@@

In general cultural distance does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.

>>>>

The overall picture is that during the Bronze Age in Europe there was a trend away from body burial towards cremation. Around 2200 calBC cremation burials were almost unknown in Poland, Southern Germany, Northeastern France and Northern Germany. In the Northwest, in Great Britain and the Benelux area, and in the Southeast, in Austria and the Czech Republic, however, there were early cremation traditions. These contexts could be discussed as areas of origin of the later omnipresent phenomenon. While cremation burials lost importance in Austria and the Czech Republic as well as in the Benelux region during the Early Bronze Age and only regained relevance there in the Middle Bronze Age, their share increased steadily in England. Inhumation and cremation were balanced here for a long time. The situation was similar in Southern Scandinavia, where the proportion of cremation burials was much lower until the Late Bronze Age. In Poland and Southern Germany, cremation became relatively suddenly and completely established with the beginning of the Late Bronze Age. In Northeastern France and Northern Germany this change took place already in the Middle Bronze Age. In both contexts inhumation continued to play an important role.

@@

Table \ref{tab:dactb} shows that the distance networks of *burial type* and *burial construction* are not equivalent. There is also no significant negative correlation. The degree of agreement between two regions in the question of inhumation and cremation is therefore not suitable for predicting the similarity in the dominance of flat or mound graves. Only for the time slot at the transition to the Middle Bronze Age from 1800 to 1600 calBC the 5% significance level of the Mantel test is almost undercut, thus suggesting a statistically significant relationship. This moment of strong correlation is though not induced by similarity along the spatial dimension: The effect is still visible in the result of the Partial Mantel test, where the spatial distance is artificially counterbalanced (see also Figure \ref{correlationtimeseries}).

@@

For the variable *burial type* an exception with significant correlation between the distance measures only reveals itself for the Early Bronze Age time slot 2000 to 1800 calBC. In the Early Bronze Age, long before the universal switch to the Urnfield burial tradition, large regions directly adjacent to each other tend to see the same developments. In the Middle Bronze Age the SED developments of the *burial type* variable undergo the peaks already observed. The correlation between cultural and spatial distance during these peaks is low and even tilts between 1600 and 1400 calBC into the negative -- but not significantly. The cultural distance between several geographically distant regions is then smaller than between several directly adjacent regions.

@@

In general, cultural distance defined here does not seem to correlate strongly with spatial distance, especially in the Middle and Late Bronze Age.

### 2019-03-31 article/04_Discussion.Rmd
more phrasing and minor improvements

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} & \ref{tab:dactB} & Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data. The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:

>>>>

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data. The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:
